### 2025-08-28 docs/thesis_manuscript/manuscript.tex
now with custom starting genotype/genome!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LaTeX Template: Project Titlepage
%
% Source: http://www.howtotex.com
% Date: April 2011
%
% This is a title page template which be used for articles & reports.
%
% Feel free to distribute this example, but please keep the referral
% to howtotex.com
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% --------------------------------------------------------------------
% Preamble
% --------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt,twoside]{scrartcl}       % KOMA

\usepackage[a4paper,pdftex]{geometry}   % A4paper margins
\setlength{\oddsidemargin}{5mm}                 % Remove 'twosided' indentation
\setlength{\evensidemargin}{5mm}

\usepackage[english]{babel}
\usepackage[font=footnotesize, labelfont=it]{caption}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subfig}
\usepackage{amsfonts}
\usepackage[backref=true, sorting=none]{biblatex}
\usepackage{csquotes}
\addbibresource{bibliography.bib}
\usepackage{hyperref}
\hypersetup{
        colorlinks=true,
        linkcolor=blue,
        filecolor=magenta,
        urlcolor=blue,
        citecolor=blue
        }
\urlstyle{same}

% --------------------------------------------------------------------
% Definitions (do not change this)
% --------------------------------------------------------------------

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}   % Horizontal rule

\makeatletter                                                   % Title
\def\printtitle{%
    {\centering \@title\par}}
\makeatother

\makeatletter                                                   % Author
\def\printauthor{%
    {\centering \large \@author}}
\makeatother

% --------------------------------------------------------------------
% Metadata (Change this)
% --------------------------------------------------------------------

\title{
                        \HRule{0.5pt} \\                                                % Upper rule
                        \LARGE \textbf{\uppercase{A Computationally Tractable Model for the Evolution of a Genome's Size and its Growing Fitness Landscape}}      % Title
                        \HRule{2pt} \\ [0.5cm]          % Lower rule + 0.5cm spacing
                        \normalsize \today                      % Todays date
                }

\author{
                \textbf{Master's Thesis}\\
                Faculty of Science, University of Bern\\
                \vspace{1cm}
                Handed in by \\
                \textbf{Jacob Hugh Riina} \\
                \vfill
                Supervisors\\
                Prof. Dr. Claudia Bank\\
                Dr. Stephan Peischl\\
}

\begin{document}

% ------------------------------------------------------------------------------
% Maketitle
% ------------------------------------------------------------------------------

\thispagestyle{empty}           % Remove page numbering on this page

\printtitle                                     % Print the title data as defined above
        \vfill
\printauthor                            % Print the author data as defined above
\newpage

% ------------------------------------------------------------------------------
% Begin document
% ------------------------------------------------------------------------------

\setcounter{page}{1}            % Set page numbering to begin on this page

\section*{Introduction}

\subsection*{Genome Evolution}

Since life's origins approximately 3.5 billion years ago, it has generally increased in complexity, and understanding the processes that generate this complexity is one of the main goals of evolutionary biology. One such mechanism of generating complexity is through genome evolution: by changing the size of their genomes, the earliest living cells - those with a cell membrane, which encode and pass on information through their DNA - were able to create \textit{evolutionary innovations}: new metabolism, physiology or regulation, which augumented their ability to thrive in a variety of environments. At the increased cost of replicating and proofreading more DNA, organisms have since been able to evolve new genes allowing them to metabolize different carbon, nitrogen, or complex nutrient sources, various types of receptors and secretion systems to interact, competitively or mutualistically, with others, and the internal regulatory networks required for them to decide which behaviors are optimal under which circumstances, or even to coordinate multicellularity. Several billion years later, we see the consequences of this incessant evolutionary process: the stunning diversity amongst organisms in the tree of life, each a product of the series of environments they have adapted to. We can say, then, that the content of an organism's genome is not solely the result of how it has evolved to fit its current environment, but also the entire history of environments that its ancestors have evolved in; as always, "Nothing in biology makes sense except in the light of evolution".\\

Understanding this facet of evolution is [difficult]. Unlike with [allelic mutation], genome evolution occurs up to several orders of magnitude less freqently, and is more heterogenous in its manifestation. We define here genome evolution as a mutational process resulting in either the expansion or reduction in the size of the [functional/coding] genome, stemming from one of several causes: amongst genome-increasing mechanisms, there exist duplications from repeat region replication errors, transposon activity, or whole genome duplications, as well as genes arising from noncoding sequence through de novo gene birth, while genome-reducing mechanisms are pseudogenization and deletions via various replication errors. here are some more specifics about the expansion/reduction process. here is why they are difficult to observe or create experiments to test. here is why is would be beneficial to create models that allow us to ask [] questions that bridge the gap between theory and experiment. 
Owing to a large protein fold space, evolution has generally favored mechanisms of duplication and neofunctionalization amongst new genes through horizontal gene transfer, transposon activity, or whole-genome duplications. (by doing so the cell acheives two things: 1. relieve pleiotropic effects, 2. lessen time to navigate sequence space). Alternatively, unused regions of the geome may also see genes spontaneously arise. (this process is much slower and harder to observe). In either case, these new genes are expected to accumulate neutrally, and over time have the opportunity to gain beneficial functionality. 
Although organisms are effective at removing unnecessary genes through pseudogenization, (this is a slow process, and still does not account for historical contingencies)
Different organisms may happen to adapt to the same environment in different way

We do however see in even shorter evolution experiments that not only mutation of alleles, but also genomic abberations appear that explain increases in fitness. (noted rate of nucleotide gain per generation stats here)
\subsection*{Fitness Landscape Theory}

As has been noted by studies of the \textit{de novo} evolution of genes, their emergence out of unused genetic substrate remains difficult to observe experimentally. (why? duplication vs de novo navigation of landscapes). There is thus a usefulness to generation of models that allow us to form hypotheses about these difficult to test systems.

Although fitness landscape theory has been used extensively to model adaptation on static genomes, it is less well developed on the scale of long term evolution, where changes in the genome size also arise. With this work we extend the Rough Mount Fuji model to incorporate genome evolution alongside allelic mutation

\section*{Methods}

\subsection*{The Rough Mount Fuji Model}

The Rough Mount Fuji (RMF) model allows for tunable epistasis \cite{neidhartAdaptationTunablyRugged2014} and so enables the investigation its effects as landscapes scale. In this model we represent genotypes as binary vectors, and their permutations in genotype space construct a network where each has a hamming distance of 1 from its neighbors, and the degree of each node is equivalent to the number of loci $L$ in the model (see Figure~\ref{fig:genotype_space}).

The assignment of fitness values to genotypes entails sampling two probabilistically determined components: Locus-wise additive effects $a = (a_1, \cdots, a_i) \sim \text{iid. } \mathcal{N}(0, 1)$ are drawn from a normal distribution, and each genotype $g$ recieves a unique epistatic component $e_g \sim \mathcal{N}(0, \sigma_e)$, also drawn from a normal distribution.  Fitness values for some arbitrary genotype can then be assigned as the sum of the product of the additive effects and the genotype, plus the epistatic component, as an exponential: $$f_g = exp{(e_g + \sum_{i=0}^{L} a_i * g_i)}$$ Epistasis is represented in the ratio between the variance of additive components and epistatic components, and we control epistasis on the landscape by varying the epistatic $\sigma_e$ - in cases where the epistatic component is small the additive component dominates and the landscape becomes easily predictable, whereas when epistatic effects are large and outweigh the additive component, there is virtually no correlation between different genotypes and ruggedness is high. Now able to reproducibly retrieve some epistatic component for any genotype by providing the CBRNG with the base-10 representation of that binary genotype, we are no longer obligated to generate fitnesses for all genotypes on the landscape in advance and instead need only do so upon encountering that genotype during simulation.

A discrete-generation RMF model with Wright-Fisher dynamics was implemented based on \cite{liRapidAdaptationRecombining2024} and previous software \cite{amadoSTUNForwardtimeSimulation2023}. We model a static $N$ total individuals mutating and reproducing simultaneously in each generation, with genotype frequencies governed by binomial and multinomial sampling, respectively. To initialize the simulation, we specify the number of loci in the model and consequently the length of the vector of additive effects to generate, the number of randomly distributed initial genotypes on the landscape and the number of individuals per genotype, the simulation length, the variance $\sigma$ of the epistatic component, and the mutation rate $\mu$ - the probability for an individual to generate a mutant of hamming distance 1 from its parent. An initial report table is then generated with a description of initial genotypes on the landscape, their population sizes, and fitnesses. We proceed to instantiate subsequent generations by first generating mutants from each extant genotype with population $\mathbb{P}$ through binomial sampling: $\text{Binomial}(n = \mathbb{P}, p = \mu)$. We adjust the corresponding population counts and calculate fitnesses for new genotypes if necessary, and then assign from $i$ extant genotypes for all individuals in the new generation of size $N$ new genotypes according to relative fitness: $\text{Multinomial}(N, p^{\prime} = (p^{\prime}_1, \cdots, p^{\prime}_i))$ given $p^{\prime}_i = p_i * \frac{\omega_i}{\bar{\omega}}$, where $p_i$ is some genotype's current frequency, $\omega_i$ is its fitness, and $\bar{\omega}$ is the average fitness across all extant genotypes. Therefore, highly fit mutants gain an advantage during sampling and are less likely to be eliminated through drift. The simulation is run for the specified number of generations and upon terminating a final report is also provided (Figure~\ref{fig:report}).

\subsection*{Counter-Based Random Number Generation}

The facile generation of random epistatic components was performed using a Julia implementation \cite{JuliaRandomRandom123jl2024} of the Random123 package from D.E. Shaw Research \cite{salmonParallelRandomNumbers2011}. Particularly, the ARS1x generator was used, with a period of $2^{128}$, this being the number of loci that can be modeled in a genotype network without reseeding.

\subsection*{Scripting}

\section*{Results}

\subsection*{Additive Limit}

\subsection*{Uncorrelated Limit}

\subsection*{Time-Dynamic Outputs}

\subsection*{Endpoint Outputs}

Does endpoint exist under this model?

\section*{Discussion}

\section*{Code and Data Availability}

\printbibliography

\end{document}